openarchaeoCollaboration

### 2021-06-15 analysis/paper.Rmd
Begin generating figures for CAA2021 slides

@@

```{r plot-categories}
oarch %>%
  drop_na(category) %>%
  group_by(category) %>%
  summarise(n = n()) %>%
  ggplot(aes(fct_reorder(category, n), n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most popular categories",
    subtitle = "By number of projects",
    caption = "Data: open-archaeo.info",
    x = NULL, y = NULL
  )
```

```{r plot-platforms}
oarch %>%
  drop_na(platform) %>%
  group_by(platform, author1_name) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(total = sum(n)) %>% 
  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
  geom_col(colour = "white") +
  coord_flip() +
  labs(
    title = "Most popular platforms",
    subtitle = "By number of projects, grouped by primary author",
    caption = "Data: open-archaeo.info",
    x = NULL, y = NULL
  )
```

@@

oarch %>% 
  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
               names_to = "repo_host", values_to = "repo_url") %>% 
  drop_na(repo_url) %>% 
  group_by(repo_host, author1_name) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  mutate(repo_host = fct_reorder(repo_host, total),
         repo_host = fct_relevel(repo_host, "none")) %>% 
  ggplot(aes(repo_host, n, group = author1_name)) +
  geom_col(colour = "#333333") +
  coord_flip() +
  labs(title = "Most popular repository hosts",
       subtitle = "By number of projects, grouped by first author",
       caption = "Data: open-archaeo.info",
       x = NULL, y = NULL)
```

@@

```{r plot-gh_commits-cum}
oarch %>%
  unnest(gh_commits) %>%
  drop_na(sha) %>%
  mutate(date = floor_date(datetime, "week")) %>%
  mutate(lumped_category = recode(category,
                                  "Specifications, protocols and schemas" = "Documents",
                                  "Products" = "Documents",
                                  "Guides" = "Documents",
                                  "Lists and datasets" = "Documents",
                                  .default = category)) %>% 
  group_by(lumped_category, date) %>%
  summarise(n_commits = n(), .groups = "drop_last") %>%
  mutate(cat_total = sum(n_commits)) %>%
  mutate(cum_commits = cumsum(n_commits)) %>%
  ggplot(aes(date, cum_commits,
             fill = fct_reorder(lumped_category, cat_total),
             colour = fct_reorder(lumped_category, cat_total))) +
  geom_area() +
  scale_x_datetime(limits = c(ymd("2008-01-01", tz = "UTC"), NA),
                   date_breaks = "1 years",
                   date_labels = "%Y") +
  scale_colour_bright(guide = NULL) +
  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Growth of open archaeology",
    subtitle = "Measured by cumulative commits in GitHub repositories",
    caption = "Data: <https://open-archaeo.info> and the GitHub API",
    x = NULL, y = NULL, fill = NULL
  )
```

@@

```{r plot-lifespan}
min_commits <- 5

oarch %>% 
  unnest(gh_commits) %>% 
  drop_na(datetime) %>% 
  group_by(item_name) %>%
  add_tally() %>% 
  filter(n >= min_commits) %>% 
  mutate(first_commit = min(datetime),
         mid_commit = median(datetime),
         last_commit = max(datetime)) %>%
  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = TRUE),
             y = datetime,
             ymin = first_commit,
             ymax = last_commit)) +
  geom_linerange() +
  geom_point(shape = "|") +
  scale_y_datetime(date_breaks = "1 year") +
  coord_flip() +
  labs(
    title = "Project lifespan",
    subtitle = glue("Individual commits in GitHub repositories (with {min_commits} or more commits)"),
    x = NULL, y = NULL,
    caption = "Data: open-archaeo.info and GitHub API"
  )
```

@@

```{r plot-gh_contributors-histogram}
oarch %>%
  drop_na(gh_repo) %>%
  unnest(gh_contribs) %>%
  group_by(gh_repo) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = n)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    title = "Repository contributors",
    caption = "Data: <https://open-archaeo.info> and the GitHub API",
    x = "Number of contributors", y = "Number of repositories"
  )
```

>>>>

```{r plot-theme}
theme_oarch <- function() theme(
  text = element_text(family = "Lato")
)
```

@@

```{r plot-categories}
oarch <- mutate(oarch, 
                category = recode(category,
                                  "Specifications, protocols and schemas" = "Specifications etc.",
                                  .default = category))
oarch %>%
  drop_na(category) %>%
  group_by(category) %>%
  summarise(n = n()) %>%
  ggplot(aes(fct_reorder(category, n), n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most popular categories",
    subtitle = "By number of projects",
    caption = "Data: open-archaeo.info",
    x = NULL, y = NULL
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
```

```{r plot-platforms}
oarch %>%
  drop_na(platform) %>%
  group_by(platform, author1_name) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(total = sum(n)) %>% 
  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
  geom_col(colour = "#333333") +
  coord_flip() +
  labs(
    title = "Most popular platforms",
    subtitle = "By number of projects, split by primary author",
    caption = "Data: open-archaeo.info",
    x = NULL, y = NULL
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
```

@@

oarch %>% 
  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
               names_to = "repo_host", values_to = "repo_url") %>% 
  drop_na(repo_url) %>% 
  group_by(repo_host, author1_name) %>% 
  summarise(n = n(), .groups = "drop_last") %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  mutate(repo_host = fct_reorder(repo_host, total),
         repo_host = fct_relevel(repo_host, "none")) %>% 
  ggplot(aes(repo_host, n, group = author1_name)) +
  geom_col(colour = "#333333") +
  coord_flip() +
  labs(title = "Most popular repository hosts",
       subtitle = "By number of projects, split by primary author",
       caption = "Data: open-archaeo.info",
       x = NULL, y = NULL)
```

@@

```{r plot-gh_commits-cum}
oarch %>%
  unnest(gh_commits) %>%
  drop_na(sha) %>%
  mutate(date = floor_date(datetime, "week")) %>%
  mutate(lumped_category = recode(category,
                                  "Specifications etc." = "Documents",
                                  "Products" = "Documents",
                                  "Guides" = "Documents",
                                  "Lists and datasets" = "Documents",
                                  .default = category)) %>% 
  group_by(lumped_category, date) %>%
  summarise(n_commits = n(), .groups = "drop_last") %>%
  mutate(cat_total = sum(n_commits)) %>%
  mutate(cum_commits = cumsum(n_commits)) %>%
  ggplot(aes(date, cum_commits,
             fill = fct_reorder(lumped_category, cat_total),
             colour = fct_reorder(lumped_category, cat_total))) +
  geom_area() +
  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"), 
                              ymd("2021-05-31", tz = "UTC")),
                   date_breaks = "2 years",
                   date_labels = "%Y") +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_bright(guide = NULL) +
  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Growth of open archaeology",
    subtitle = "Measured by cumulative commits in GitHub repositories",
    caption = "Data: open-archaeo.info and GitHub API",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_oarch()

# Export for slides (full slide)
# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
```

@@

```{r plot-lifespan}
oarch %>% 
  unnest(gh_commits) %>% 
  drop_na(datetime) %>% 
  group_by(item_name) %>%
  add_tally() %>% 
  filter(n > 1) %>% 
  mutate(first_commit = min(datetime),
         mid_commit = median(datetime),
         last_commit = max(datetime)) %>%
  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
             y = datetime,
             ymin = first_commit,
             ymax = last_commit)) +
  geom_linerange(size = 0.05) +
  geom_point(size = 0.1) +
  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"), 
                              ymd("2021-05-31", tz = "UTC")),
                   date_breaks = "2 years", date_labels = "%Y") +
  # coord_flip() +
  labs(
    title = "Project lifespan",
    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
    x = NULL, y = NULL,
    caption = "Data: open-archaeo.info and GitHub API"
  ) +
  theme_oarch() +
  theme(
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank()
  )

# Export for slides (full slide)
# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
```

@@

```{r plot-gh_contributors-histogram}
oarch %>%
  drop_na(gh_repo) %>%
  unnest(gh_contribs) %>%
  group_by(gh_repo) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = n)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    title = "Repository contributors",
    caption = "Data: <https://open-archaeo.info> and the GitHub API",
    x = "Number of contributors", y = "Number of repositories"
  )

```
